\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\geometry{margin=1in}

\title{Pintle Engine Simulation Pipeline:\\
Physics, Algorithms, and Software Workflow}
\author{Kush Mahajan}
\date{\today}

\begin{document}

\maketitle
\tableofcontents
\newpage

\section{Purpose and Scope}

This document summarizes the physical models, numerical algorithms, and software workflow implemented in the pintle engine simulation suite.  It covers all core packages:
\begin{itemize}
    \item \verb|pintle_pipeline| -- shared physics (thermochemistry, feed dynamics, cooling, configuration)
    \item \verb|pintle_models| -- injector solvers, chamber closure, nozzle performance
    \item \verb|examples/pintle_engine| -- orchestration scripts, plotting, validation tooling
\end{itemize}
The pipeline converts tank pressure time histories into chamber pressure, mixture ratio, cooling diagnostics, and thrust, with extensibility for multiple injector families and cooling strategies.

\section{Software Architecture Overview}

\subsection{Module Map}
\begin{itemize}
    \item \textbf{Configuration schemas} (\verb|config_schemas.py|): strongly-typed inputs for all subsystems using Pydantic.
    \item \textbf{Thermochemistry} (\verb|cea_cache.py|, \verb|combustion_eff.py|): RocketCEA precomputation with bilinear interpolation, combustion-efficiency corrections.
    \item \textbf{Feed \& cooling} (\verb|feed_loss.py|, \verb|regen_cooling.py|, \verb|film_cooling.py|, \verb|ablative_cooling.py|): hydraulic losses and thermal management.
    \item \textbf{Injector models} (\verb|injectors/*.py|): pintle, shear-coaxial, and impinging implementations with shared closure logic.
    \item \textbf{Chamber closure} (\verb|chamber_solver.py|): supply vs.\ demand balance solved via Brent's method.
    \item \textbf{Nozzle dynamics} (\verb|nozzle.py|): thrust coefficient and exit properties with finite-efficiency correction.
    \item \textbf{Runner} (\verb|runner.py|): end-to-end orchestration, diagnostics, batch evaluation.
    \item \textbf{Interactive / batch workflows} (\verb|examples/pintle_engine/*|): UI, plotting, validation harnesses.
\end{itemize}

\subsection{Data Flow}
At each operating point the following order is enforced:
\begin{enumerate}
    \item Read configuration $\rightarrow$ instantiate \verb|PintleEngineRunner|.
    \item For given tank pressures $(P_{\text{tank},O}, P_{\text{tank},F})$, call the chamber solver.
    \item Injector closure returns mass-flows $\dot{m}_O, \dot{m}_F$ and spray diagnostics.
    \item Thermochemical properties from the CEA cache adjust combustion efficiency.
    \item Residual $R = \dot{m}_{\text{supply}} - \dot{m}_{\text{demand}}$ drives Brent root-finding to converge $P_c$.
    \item Once converged, nozzle model and cooling modules compute thrust and wall conditions.
\end{enumerate}

\section{Configuration Schema}

\subsection{Fluids and Materials}
Each fluid (\verb|FluidConfig|) stores density, viscosity, surface tension, vapor pressure, specific heat, thermal conductivity, and bulk temperature.  Cooling models reference these for heat-transfer correlations.

\subsection{Injector Families}
Three injector classes comply with the interface in \verb|injectors/__init__.py|:
\begin{itemize}
    \item \textbf{Pintle}: axial LOX through drilled orifices, radial fuel through an annulus, configurable via \verb|PintleInjectorConfig|.
    \item \textbf{Coaxial}: core oxidizer jets with annular fuel and optional swirl, defined by \verb|CoaxialInjectorConfig|.
    \item \textbf{Impinging}: paired jet elements producing an impact sheet, specified by \verb|ImpingingInjectorConfig|.
\end{itemize}
Each configuration supplies geometry, optional swirl/impingement angles, and is validated during construction.

\subsection{Feed Systems}
Feed branches (\verb|FeedSystemConfig|) define diameter (or hydraulic area) and loss correlation coefficients $(K_0, K_1)$ with pressure response $\phi(P)$ selected from \texttt{none}, \texttt{sqrtP}, or \texttt{logP}.  These parameters feed the generalized loss model in Section~\ref{sec:feed}.

\subsection{Cooling Blocks}
\begin{itemize}
    \item \verb|RegenCoolingConfig|: channel geometry, roughness, manifold losses, optional conjugate heat-transfer.
    \item \verb|FilmCoolingConfig|: mass-fraction allocation, effectiveness decay, optional injection temperature.
    \item \verb|AblativeCoolingConfig|: thermo-physical properties of ablative liners for terminal protection.
\end{itemize}

\subsection{Combustion and Solver}
\verb|CombustionConfig| holds RocketCEA settings and efficiency model parameters.  \verb|SolverConfig| specifies chamber-pressure bounds, method (Brent, secant, Newton), tolerances, and nested closure options (max iterations, $C_d$ reduction factor).

\section{Thermochemistry and CEA Cache}

\subsection{Cache Construction}
\verb|cea_cache.py| precomputes thermo-chemical properties over a structured grid:
\[
P_c \in [P_{\min}, P_{\max}], \qquad \text{MR} \in [\text{MR}_{\min}, \text{MR}_{\max}],
\]
with $n_{\text{points}}^2$ samples.  For each grid point, RocketCEA returns $c^*_{\text{ideal}}$, $C_f$, $T_c$, $\gamma$, $R$, $M$, and sea-level $I_{sp}$ proxies.  Results are serialized into \verb|.npz| with metadata for reproducibility; mismatched metadata triggers automatic rebuilds.

\paragraph{Physical Inputs.}  RocketCEA is configured with oxidizer/fuel identifiers, mixture-ratio range, chamber-pressure range, and expansion ratio drawn from \verb|CEAConfig|.  Pressures passed to RocketCEA are converted from pascals to psia (.obj requirement) and converted back to SI to maintain consistency everywhere else.  Propellant properties assume chemical equilibrium in the chamber with frozen composition through the nozzle, aligning with the single-phase hot-gas assumption used throughout the pipeline.

\paragraph{Fallback Handling.}  If RocketCEA fails to converge at a grid point (for example at the extreme MR or $P_c$ limits) the tables store NaN.  During interpolation the solver falls back to nearest-neighbor values, preserving continuity while warning through console output so users know the cache encountered extrapolated points.

\subsection{Bilinear Interpolation}
During runtime the cache supplies properties via bilinear interpolation in $(P_c, \text{MR})$:
\[
f(P_c, \text{MR}) = \sum_{i=0}^1\sum_{j=0}^1 w_{ij} f_{ij}, \qquad w_{ij} = \frac{(1 - \xi_i)(1 - \eta_j)}{1},
\]
where $\xi$ and $\eta$ are normalized distances between bounding grid points.  NaN tiles fall back to nearest-neighbor values for robustness.

\paragraph{Out-of-Grid Behavior.}  Queries outside $[P_{\min}, P_{\max}]$ or $[\text{MR}_{\min}, \text{MR}_{\max}]$ are clipped to the closest in-bounds node before interpolation.  This zero-slope extrapolation keeps the solver stable while emitting diagnostics in the runner so operators know when operating conditions exceed cache coverage.

\paragraph{Thermodynamic Consistency.}  All interpolated quantities originate from the same RocketCEA state, preventing combinations such as $c^*$ from one MR value and $\gamma$ from another.  This guarantees that nozzle and combustion calculations downstream remain thermodynamically consistent.

\subsection{Combustion Efficiency}
The finite chamber correction \verb|eta_cstar| implements:
\[
\eta_{c^*} =
\begin{cases}
1 - C & \text{constant}, \\
1 - C e^{-K L^*} & \text{exponential (default)}, \\
1 - C \left(1 - \frac{L^*}{1~\text{m}}\right) & \text{linear},
\end{cases}
\]
optionally penalized if spray constraints fail.  Actual characteristic velocity follows $c^*_{\text{actual}} = \eta_{c^*} c^*_{\text{ideal}}$, leading to chamber demand in Section~\ref{sec:chamber}.

\paragraph{Finite-Rate Interpretation.}  The exponential model mimics finite residence-time data: short chambers (small $L^*$) lose efficiency; large $L^*$ asymptotically approach the CEA equilibrium limit.  When spray penalties are enabled the solver multiplies $\eta_{c^*}$ by a factor $<1$ whenever atomization quality is poor, capturing the impact of incomplete mixing on burn efficiency.

\paragraph{Derived Thermodynamic States.}  Helper routines convert $\eta_{c^*}$ into an adjusted chamber temperature,
\[
T_{c,\text{actual}} = T_{c,\text{ideal}} \frac{\eta_{c^*}}{1 - (1 - \eta_{c^*})(\gamma - 1)/\gamma},
\]
which feeds the cooling and nozzle models.  The ratio-of-specific-heats correction (\verb|calculate_frozen_flow_correction|) similarly adjusts $\gamma$ to account for frozen-flow effects in the divergent section.

\section{Feed and Distribution Losses}
\label{sec:feed}

\subsection{Generalized Loss Model}
For branch mass flow $\dot{m}$ and fluid density $\rho$:
\[
\Delta p_{\text{feed}} = K_{\text{eff}}(P_{\text{tank}}) \frac{\rho}{2}
\left(\frac{\dot{m}}{\rho A_{\text{hyd}}}\right)^2,
\quad
K_{\text{eff}} = K_0 + K_1 \phi(P_{\text{tank}}).
\]
Hydraulic area defaults to the circular inlet diameter when omitted.  The model supports pressure-dependent valves or regulators via $\phi(P)$.

\paragraph{Flow-Regime Awareness.}  Because the loss model operates on velocity, it seamlessly spans laminar and turbulent regimes.  $K_0$ captures fixed geometric losses (elbows, screens) while $K_1$ and $\phi(P)$ represent valves or regulators whose effective loss coefficient varies with pressure.  Choosing $\phi(P) = \sqrt{P}$ approximates choked flow regulators; $\phi(P) = \log P$ fits compressible orifices.

\paragraph{Hydraulic Area Resolution.}  If \verb|A_hydraulic| is omitted, the code derives it from the specified \verb|d_inlet|, ensuring feed velocities remain consistent even when users provide only diameters.

\subsection{Regenerative Channels}
\verb|delta_p_regen_channels| decomposes the coolant path into inlet pipe, split manifold, $N$ parallel rectangular channels, merge and outlet pipe.  Each segment accounts for friction ($f$ via Blasius or Swamee--Jain) and entrance/exit minor losses based on Reynolds-adjusted discharge coefficients:
\[
C_d(Re) = C_{d,\infty} - \frac{a_{\text{Re}}}{\sqrt{Re}}\quad \text{clamped to }[C_{d,\min}, C_{d,\infty}].
\]
Channel hydraulic diameter is $D_h = 2wh/(w+h)$.  Pressure drops accumulate to inform injector inlet pressure on the fuel branch.

\paragraph{Segment Physics.}
\begin{itemize}
    \item \textbf{Inlet/Outlet Pipes:} Darcy--Weisbach with $f = 0.316\,\text{Re}^{-0.25}$ (Blasius) for turbulent smooth pipes or Swamee--Jain when roughness is provided.
    \item \textbf{Manifolds:} Configurable loss coefficients model sudden expansions/contractions at split and merge junctions.
    \item \textbf{Entrance/Exit Losses:} Dynamic $C_d$ terms mimic flow contraction into and expansion out of micro-channels.
    \item \textbf{Channel Core:} Rectangular-duct friction accounts for viscous shear along the hot wall over the channel length, typically equal to chamber length.
\end{itemize}

\paragraph{Mass Partitioning.}  The solver splits coolant mass flow equally across $N$ channels under the identical-geometry assumption.  Diagnostics report per-channel mass flow, velocities, and Reynolds numbers so users can verify turbulent operation or evaluate single-channel heat flux.

\section{Injector Models and Closure}

\subsection{Shared Closure Loop}
All injectors derive from the abstract \verb|InjectorModel| and implement a fixed-point iteration:
\begin{enumerate}
    \item Initialize $\dot{m}_O$, $\dot{m}_F$, and effective discharge coefficients $C_{d,\text{eff}}$.
    \item Iterate feed pressure losses (including regen contribution) to determine injector inlet pressures.
    \item Update $C_d$ from Reynolds number, pressure, and temperature.
    \item Compute new mass flows using
    \[
        \dot{m} = C_d A \sqrt{2 \rho \Delta p_{\text{inj}}}, \quad \Delta p_{\text{inj}} = P_{\text{inj}} - P_c.
    \]
    \item Evaluate spray metrics; if constraints fail, reduce $C_{d,\text{eff}}$ by the configured factor and repeat.
\end{enumerate}
Diagnostics include iteration count, $J$, TMR, Weber numbers, SMD, and constraint violations.

\subsection{Discharge Coefficient Physics}
\label{sec:cd}

Each injector element is modeled as an orifice with a discharge coefficient that captures the departure from ideal, inviscid flow.  The base expression is
\[
\dot{m} = C_d \, A \sqrt{2 \rho \Delta p_{\text{inj}}},
\qquad \Delta p_{\text{inj}} = P_{\text{inj}} - P_c,
\]
where $A$ is geometric area and $\rho$ is inlet density.  The code converts this single $C_d$ into a compound model that accounts for Reynolds-number scaling, compressibility, and temperature-dependent viscosity effects.

\paragraph{Reynolds Scaling.}  For each branch the instantaneous Reynolds number is
\[
\text{Re} = \frac{\rho u D_h}{\mu}, \qquad u = \frac{\dot{m}}{\rho A}.
\]
The empirical form
\[
C_d^{\text{Re}} = C_{d,\infty} - \frac{a_{\text{Re}}}{\sqrt{\text{Re}}}
\]
matches classical sharp-edged-orifice data, where $C_{d,\infty}$ represents the asymptotic turbulent value and $a_{\text{Re}}$ tunes low-Re losses.  For vanishing $\text{Re}$ the solver clamps $C_d$ to a configured minimum $C_{d,\min}$ to avoid singularities.

\paragraph{Pressure (Compressibility) Correction.}  At high inlet pressures, density increase and vena contracta changes are modeled through
\[
C_d^{P} = C_d^{\text{Re}} \left[ 1 + a_P \left(\frac{P_{\text{inj}}}{P_{\text{ref}}} - 1 \right) \right],
\]
with $a_P$ typically small and $P_{\text{ref}}$ a representative pressure (default $5\ \text{MPa}$).  Positive $a_P$ slightly raises $C_d$ as pressure rises; negative values capture compressibility-induced contraction.

\paragraph{Temperature Correction.}  Viscosity variations with inlet temperature $T_{\text{inj}}$ enter as
\[
C_d^{T} = C_d^{P} \left[ 1 + a_T \left(\frac{T_{\text{inj}}}{T_{\text{ref}}} - 1 \right) \right].
\]
Reducing viscosity (higher temperature) effectively raises Reynolds number, so $a_T > 0$ boosts $C_d$.

\paragraph{Clamping and Iteration Coupling.}  After pressure and temperature adjustments the model enforces
\[
C_d = \operatorname{clip}\bigl(C_d^{T},\, C_{d,\min},\, C_{d,\infty}\bigr)
\]
and reuses the updated value in the mass-flow equation.  Because $C_d$ depends on $\dot{m}$ (via $\text{Re}$) the closure loop iteratively recomputes flow velocity and $C_d$ until residuals fall below the configured tolerance.

\paragraph{Spray Constraint Feedback.}  When Weber or evaporation constraints fail, the solver multiplies the current $C_{d,\text{eff}}$ by a reduction factor (default $0.95$) before the next iteration.  This mimics throttling or partial flow separation that would reduce effective injector area under poor atomization, steering the iteration toward a low-flow solution that satisfies spray physics.

\paragraph{Physical Interpretation.}  The combined model emulates the dominant physics of subsonic orifices: viscous boundary layers reduce $C_d$ at low Re, compressibility effects alter the vena contracta at high pressure, and viscosity changes shift effective Reynolds number.  Clamping keeps $C_d$ within $[C_{d,\min}, C_{d,\infty}]$, preventing unphysical values and improving numerical stability.

\subsection{Pintle Injector}
The pintle implementation (\verb|injectors/pintle.py|) uses drilled LOX orifices plus an annular fuel gap.  Geometric helpers in \verb|geometry.py| compute effective areas and hydraulic diameters directly from configuration:
\[
A_{\text{LOX}} = N \frac{\pi d_{\text{or}}^2}{4}, \qquad
A_{\text{fuel}} = \pi(R_{\text{outer}}^2 - R_{\text{inner}}^2).
\]
Fuel regen pressure drop is seamlessly included when enabled.

\subsection{Shear Coaxial Injector}
The coaxial model adds annular swirl handling: axial velocity $u_{F,\text{axial}}$ is increased by $\cos\theta_{\text{swirl}}$ to estimate total jet velocity and Weber number.  Port area is $A_{\text{core}} = n\pi d^2 / 4$ and annulus area computed via difference of diameters.  Spray diagnostics incorporate the swirl angle in output.

\paragraph{Swirl Physics.}  Tangential entry augments shear at the oxidizer/fuel interface.  The model conserves mass by using axial velocity in momentum balance while attributing additional kinetic energy to swirling motion, raising relative velocity and Weber number---a first-order approximation of the enhanced atomization observed in shear-coaxial elements.

\subsection{Impinging Injector}
The impinging configuration models paired jets forming a sheet with relative velocity
\[
u_{\text{sheet}} = \sqrt{u_O^2 + u_F^2 - 2 u_O u_F \cos \theta_{\text{imp}}}.
\]
Spray correlation uses the same Weber/Ohnesorge framework, tailored by element impingement angles.

\section{Spray, Atomization, and Mixing}

\subsection{Momentum Flux Ratio}
\[
J = \frac{\rho_O u_O^2}{\rho_F u_F^2}.
\]
The thrust-momentum ratio $TMR = J / (1+\text{MR})$ drives alternative spray-angle fitting.

\paragraph{Interpretation.}  $J$ compares the axial momentum densities of oxidizer and fuel streams.  Large $J$ pulls the spray toward the axis (narrow cone), while small $J$ lets the fuel sheet dominate, producing wider plumes.  Because $J$ scales with density and velocity squared, chilled oxidizers or high-pressure drops can dramatically bias spray structure.

\subsection{Spray Angle Models}
Two selectable parameterizations are available:
\begin{align*}
\theta &= 2 \arctan\!\left( k J^n \right) & \text{(momentum flux)} \\
\theta &= \arccos\!\left( \frac{1}{1 + TMR^{0.75}} \right) & \text{(TMR-based)}.
\end{align*}

\subsection{Weber, Ohnesorge, and SMD}
Weber numbers quantify disruptive aerodynamic forces:
\[
\text{We} = \frac{\rho u^2 d}{\sigma},
\qquad
\text{Oh} = \frac{\mu}{\sqrt{\rho \sigma d}}.
\]
Lefebvre's correlation yields Sauter mean diameter
\[
D_{32} = C d \text{We}^{-m} \text{Oh}^p.
\]
Evaporation length $x^* = U_{\text{rel}} K D_{32}^2$ must remain below the configured limit to ensure complete vaporization inside the chamber.

\paragraph{Flow-Property Coupling.}  Fluid properties $\rho$, $\mu$, and $\sigma$ come directly from the configuration, so Weber and Ohnesorge numbers respond to cryogenic density increases or viscosity changes.  The solver evaluates $D_{32}$ separately for oxidizer and fuel streams, then enforces the worst-case (largest) evaporation length to guarantee both propellants vaporize before the throat.

\section{Cooling Models}
\label{sec:cooling}

Cooling strategies are configured to be mutually exclusive: at any operating point the simulation enables either film cooling, regenerative cooling, or ablative protection.  The following subsections document each model individually.

\subsection{Film Cooling}
When film cooling is selected, a fraction of the fuel stream is diverted to form a protective boundary layer along the chamber wall:
\[
\dot{m}_{\text{film}} = f_{\text{film}} \dot{m}_F,
\qquad
\dot{m}_{\text{core}} = \dot{m}_F - \dot{m}_{\text{film}}.
\]
Effectiveness decays exponentially across the covered portion of the chamber, producing a heat-flux reduction factor applied directly to the hot-gas boundary condition.

\paragraph{Temperature Mixing.}  The module forms an effective gas temperature at the wall,
\[
T_g^{\text{eff}} = T_c - \eta_{\text{film}} (T_c - T_{\text{inj}}),
\]
where $T_{\text{inj}}$ is either the specified injection temperature or the fuel bulk temperature.  This temperature propagates into the regen energy balance and ablative model.

\paragraph{Axial Coverage.}  The parameter \verb|apply_to_fraction_of_length| scales the baseline effectiveness to mimic partial axial coverage (e.g.\ film only near the injector face).  Values $<1$ reduce effectiveness accordingly, while values $>1$ allow modest over-coverage for staged injection concepts.  Because cooling models are mutually exclusive, film effectiveness does not feed any regenerative or ablative calculation.

\subsection{Regenerative Heat Transfer}
\verb|compute_regen_heat_transfer| resolves coolant-side Nusselt number (Dittus--Boelter for turbulent flow) and hot-gas coefficient using axial Reynolds number inside the chamber.  Overall heat flux combines convection and radiation:
\[
q'' = U (T_g^{\text{eff}} - T_{\text{bulk}}) + \varepsilon \sigma (T_g^{4} - T_{\text{bulk}}^{4}),
\]
with $U$ computed from series resistances of gas film, wall conduction, and coolant film.  The module returns hot/cold wall temperatures, cooling effectiveness, and outlet coolant temperature.

\paragraph{Hot-Gas Film.}  Gas density follows $\rho_g = P_c/(R_g T_c)$, and bulk velocity is $\dot{m}_{\text{total}}/(\rho_g A_{\text{chamber}})$.  The solver switches between laminar ($Nu=4.36$) and turbulent (Dittus--Boelter) regimes based on Reynolds number, using a configurable Prandtl number representative of combustion gases.

\paragraph{Coolant Film.}  Channel velocity derives from per-channel mass flow and cross-sectional area.  The same laminar/turbulent switch applies, with coolant properties (density, viscosity, conductivity, specific heat) supplied via configuration.  Iterative updates on coolant outlet temperature re-evaluate these properties to capture the effect of heating.

\paragraph{Wall Temperatures.}  Hot-wall temperature is $T_g^{\text{eff}} - q''/h_g$; cold-wall temperature subtracts the conduction term across wall thickness; coolant-interface temperature adds $q''/h_c$ to the outlet bulk temperature.  These values feed directly into ablative calculations and provide structural design inputs.

\subsection{Ablative Response}
For residual heating after regen, the ablative layer converts net heat flux into mass loss:
\[
\dot{m}'' = \frac{q''_{\text{effective}}}{H_{\text{abl}}},
\qquad
\dot{r} = \frac{\dot{m}''}{\rho_{\text{abl}}},
\]
while radiative relief reduces incoming flux when the surface approaches the temperature limit.

\paragraph{Surface Energy Balance.}  Radiative relief is computed as $\varepsilon \sigma (T_{\text{surf}}^{4} - T_{\text{limit}}^{4})$ and subtracted from the incoming flux before dividing by heat of ablation.  If the relief term exceeds the incident heat, the solver returns zero recession and zero mass flux, reflecting a cooled surface.

\paragraph{Boundary Conditions.}  In the ablative-only configuration the wall temperature estimate defaults to the chamber temperature (or any user-supplied value).  Because cooling strategies are mutually exclusive, ablative recession is evaluated without regenerative inputs; users selecting ablative protection provide heat-flux estimates appropriate for that scenario.

\section{Chamber Pressure Solver}
\label{sec:chamber}

\subsection{Residual Definition}
Supply is obtained directly from injector closure:
\[
\dot{m}_{\text{supply}}(P_c) = \dot{m}_O(P_c) + \dot{m}_F(P_c).
\]
Demand follows from choked mass-flow requirements:
\[
\dot{m}_{\text{demand}}(P_c) = \frac{P_c A_t}{c^*_{\text{actual}}(P_c,\text{MR})}.
\]
The residual
\[
R(P_c) = \dot{m}_{\text{supply}}(P_c) - \dot{m}_{\text{demand}}(P_c)
\]
is continuous and bracketed.  Brent's method (\verb|scipy.optimize.brentq|) is the default solver due to robustness; Newton iteration remains as an optional fallback.

\paragraph{Supply Curve.}  Each evaluation of $\dot{m}_{\text{supply}}$ runs the injector closure loop with the current $P_c$, so both feed losses and discharge coefficients reflect the updated pressure drop.  This produces a monotonically decreasing supply curve in most configurations, a prerequisite for stable root finding.

\paragraph{Demand Curve.}  Demand originates from the quasi-one-dimensional choked flow relation,
\[
\dot{m}_{\text{demand}} = \frac{P_c A_t}{c^*_{\text{actual}}}.
\]
Because $c^*_{\text{actual}}$ depends on $\eta_{c^*}(L^*, \text{spray})$, the demand curve captures chamber geometry and spray-induced combustion penalties.

\subsection{Bounds and Diagnostics}
Bounds default to:
\[
P_{c,\min} = \max(10^5~\text{Pa}, P_{c,\text{cfg,min}}),\quad
P_{c,\max} = \min\bigl((1-\delta)\min(P_{\text{tank},O}, P_{\text{tank},F}), P_{c,\text{cfg,max}}\bigr),
\]
with $\delta = 0.15$ providing headroom for feed losses.  Diagnostics returned from \verb|ChamberSolver.solve| include mass flows, mixture ratio, $c^*_{\text{ideal}}$, $c^*_{\text{actual}}$, $\eta_{c^*}$, $T_c$, $\gamma$, gas constant $R$, and injector closure metadata.

\paragraph{Root-Finding Safeguards.}  Prior to iteration the solver checks the residual at both bounds.  If they do not bracket a root, descriptive exceptions indicate whether supply exceeds demand everywhere (suggesting insufficient tank pressure) or vice versa (indicating the injector cannot deliver enough flow).

\paragraph{Spray-Quality Tracking.}  The closure loop records whether final spray constraints were satisfied; the chamber solver propagates this flag into diagnostics and to the combustion-efficiency penalty, making it clear when a converged operating point violates atomization limits.

\section{Nozzle Performance}

\subsection{Exit Conditions}
The nozzle model solves the supersonic branch of the area--Mach relation:
\[
\frac{A_e}{A_t} = \frac{1}{M_e}
\left[\frac{2}{\gamma+1}\left(1 + \frac{\gamma-1}{2}M_e^2\right)\right]^{\frac{\gamma+1}{2(\gamma-1)}},
\]
by Newton iteration capped at $20$ steps, clamped to $M_e \geq 1.01$.  Exit pressure is
\[
P_e = P_c \left(1 + \frac{\gamma-1}{2}M_e^2\right)^{-\gamma/(\gamma-1)},
\]
and temperature obeys the standard isentropic relation.  Exit velocity follows from energy conservation with $c_p = \gamma R / (\gamma-1)$.

\subsection{Thrust and Specific Impulse}
Momentum and pressure components yield
\[
F = \dot{m}_{\text{total}} v_e + (P_e - P_a) A_e,
\quad
I_{sp} = \frac{F}{\dot{m}_{\text{total}} g_0}.
\]
For cross-checking, the thrust coefficient method $F = C_f P_c A_t$ uses efficiency-scaled $C_f$ from the CEA cache.

\paragraph{Pressure Balance.}  During over-expanded operation the model clamps $P_e$ at ambient pressure to avoid predicting negative pressure thrust.  Users can specify altitude pressure to study vacuum or ascent cases.

\paragraph{Nozzle Efficiency.}  A single scalar efficiency multiplies $C_f$, capturing divergence losses, throat erosion, and frozen chemistry effects.  Typical values (0.96--0.99) align with hot-fire data for converging-diverging nozzles.

\section{Runner and Batch Evaluation}

\subsection{Single Operating Point}
\verb|PintleEngineRunner.evaluate| orchestrates the call chain:
\begin{enumerate}
    \item Solve chamber pressure with \verb|ChamberSolver|.
    \item Query CEA cache for thermodynamic properties.
    \item Compute thrust and exit quantities via \verb|calculate_thrust|.
    \item Allocate fuel between film and regen cooling, compute coolant diagnostics.
    \item Optional ablative response using net wall flux.
\end{enumerate}
The returned dictionary includes thrust metrics, thermochemistry, cooling breakdown, and raw diagnostics for post-processing.

\paragraph{Diagnostic Structure.}  The runner nests subsystem diagnostics: injector closure metadata (iterations, violations) and, depending on the selected cooling strategy, either film-cooling metrics, regenerative heat-transfer results, or ablative response.  Downstream plotting scripts consume these fields directly.

\subsection{Time Histories}
\verb|evaluate_arrays| vectorizes the evaluation, iterating over tank-pressure arrays and storing results in NumPy vectors while collecting solver diagnostics per time step.  Errors at individual points are caught and reported without aborting the sweep.

\paragraph{Error Handling.}  When a point fails (e.g.\ no chamber-pressure solution), the corresponding scalar entries remain NaN while the diagnostics list stores an \verb|"error"| message.  This allows trajectory post-processing to mask failed points or perform interpolation without rerunning the entire simulation.

\section{Example Workflows}

\subsection{Interactive UI}
\verb|examples/pintle_engine/ui_app.py| embeds the runner into a Streamlit-style dashboard to explore sensitivities (tank pressure sweeps, injector settings) with live plots.

\subsection{Automated Pipelines}
\verb|run_full_pipeline.py| executes prescribed scenarios (pressure sweeps, regen studies) and writes JSON outputs via \verb|pintle_pipeline.io.save_results|.  Plotting scripts (e.g.\ \verb|comprehensive_performance_plots.py|, \verb|plot_performance_summary.py|) consume these outputs for visualization.

\subsection{Validation Utilities}
Scripts such as \verb|validate_chamber_intrinsics.py| and \verb|compare_ideal_vs_actual.py| cross-check chamber physics against analytic expectations and CEA baselines, while \verb|debug_regen_iteration.py| and \verb|check_regen_impact.py| probe regenerative cooling sensitivities.

\section{Outputs and Diagnostics}

The runner's diagnostics bundle exposes:
\begin{itemize}
    \item Injector status: closure iterations, constraint violations, $C_d$ adjustments.
    \item Spray metrics: $J$, TMR, Weber numbers, SMD, evaporation length.
    \item Cooling summary: film effectiveness, regen heat flux, coolant outlet temperature, ablative mass flux.
    \item Thermochemical state: $T_c$, $\gamma$, $R$, $c^*_{\text{ideal}}$, $c^*_{\text{actual}}$, nozzle exit properties.
\end{itemize}
These fields underpin the plotting/analysis routines shipped in the examples directory.

\section{Extensibility Notes}

\begin{itemize}
    \item \textbf{Additional injectors}: implement \verb|InjectorModel.solve|, reusing utilities in \verb|spray.py| and \verb|discharge.py|.
    \item \textbf{Alternative cooling}: extend film or regen configs with new correlations; coupling in the runner isolates interface points.
    \item \textbf{Advanced chemistry}: adjust \verb|CEAConfig| to broaden MR/Pc ranges or integrate new propellant pairs by regenerating cache files.
    \item \textbf{Trajectory coupling}: \verb|evaluate_arrays| is ready for time-resolved tank models; feed in tank pressure histories and post-process thrust/Isp traces.
\end{itemize}

\section{Summary}

The implemented pipeline solves a fully coupled system in which:
\begin{itemize}
    \item Injector flow and spray physics determine supply as a nonlinear function of chamber pressure.
    \item Combustion demand depends on $c^*$ corrections tied to spray quality and chamber sizing.
    \item Cooling models feedback into available fuel mass flow and wall heat loads.
    \item Thrust is evaluated from thermochemically consistent exit conditions with nozzle efficiency.
\end{itemize}
Modular design allows each subsystem to evolve independently while the runner maintains a deterministic workflow from tank pressures to thrust and thermal diagnostics.

\section{Numerical Implementation Notes}

\begin{itemize}
    \item \textbf{Units:} All internal calculations are SI; RocketCEA calls temporarily convert to imperial units and immediately convert back.
    \item \textbf{Floors:} Small positive floors (typically $10^{-9}$) prevent divide-by-zero in Reynolds and Weber calculations when a branch starves.
    \item \textbf{Vectorization:} The array evaluator reuses the scalar path, guaranteeing identical physics whether running a single point or a full trajectory.
    \item \textbf{Caching:} Configuration objects are cached between runs in plotting scripts to avoid repeated YAML parsing; CEA cache metadata prevents accidental reuse with mismatched settings.
\end{itemize}

\section{Validation and Future Work}

\begin{itemize}
    \item Injector mass-flow predictions have been cross-checked against classical orifice equations and spreadsheet tools for representative geometries.
    \item \verb|validate_chamber_intrinsics.py| sweeps $P_c$ to verify supply-demand intersections behave monotonically and converge within tolerance.
    \item \verb|compare_ideal_vs_actual.py| contrasts ideal RocketCEA performance with efficiency-corrected outputs to validate the $c^*$ loss model.
    \item Planned extensions include two-phase film modeling, transient wall conduction for regen, and coupling to trajectory solvers for full mission analysis.
\end{itemize}

\end{document}

